### 2019-07-02 thesis-template/msc-thesis.tex
update

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/tensorfieldlines.png}
    \label{a)}
    \caption*{a) TFLs and glyphs }
  \end{minipage}
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/rings_ftle.png}
    \label{b)}
    \caption*{b) LTG($0^\circ$) for a)}
  \end{minipage}
  \caption{``rings''-test field}
\label{rings-ftle}
\end{figure}
What we can observe in Fig. \ref{rings-ftle}, is called an inside-outside test producing the ridge on the circumference of the circle with maximized radius fitting into the minimum bounding box. If one would place a light source on the circumference of this circle offsetting it spatially and in the angular domain, one would obtain one incomplete final light distribution (outside) and one complete final light distribution (inside), as this is the maximum radius at which the intensities can follow a circle. This hence leads to large differential light distributions and thus to a higher LTG magnitude. ROIs pointing outside the field respond the same for varying light source directions and positions yielding the blue areas in the corners. Regions introducing linear anisotropy respond uniformly inside of the circle in contrary to the center point, which is again not responding because of circular overlaps of the offset light distributions computed for the light transport gradient. The contour filter reveals these two edge locations in the field, concerning tensor field line divergence.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/rings_contour.png}
    \caption*{Contour}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/rings_isovolume.png}
    \caption*{IsoVolume}
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/rings_volume.png}
    \caption*{Volume}
  \end{minipage}
    \caption{LTG volume representations: ``rings'' test field}
\label{rings_contour}
\end{figure}
The volumes, we examine here in Fig. \ref{rings_contour}, show that the field is symmetric, as we do not notice a change along the $\omega\mapsto z$-axis except for the natural shift of the field against the light source direction, which we discussed already in this subsection. Along the $z$-axis, the light source is rotated in all directions which causes the shift to form a full cycle from bottom to top ($[0^\circ,360^\circ]$).
At last, we will analyze the bow test field, which is now considered to be a bit more complicated example. We again plot the FTLE field in color coding representation in Fig. \ref{bow-ftle}.

@@

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[height=\textwidth]{img/ftle_drain_alt.png}
    \caption*{LTG($0^\circ$)}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/drain_ftle_org.PNG}
     \caption*{FTLE: Hlawatsch's approach}
    \label{b)}
  \end{minipage}
\caption{\enquote{drain} test field}
\label{drain_ftle_comp}
\end{figure}
We can observe the comparison between our invented FTLE-like approach and the one of Hlawatsch et al. in the depiction in Fig. \ref{drain_ftle_comp}. Note, that a sharper but weaker ridge is detected on the right side of the image, which shines similar to a cone shape from the top. We interpret the missing spot in the top center as an artifact stemming from the way we handle our light propagation, as we allow whole distributions of trajectories to be propagated (with a scalar attribute interpretable as propability). This causes a sampling drift, we already discussed in sect. \ref{sect:total}, leading to an energy distribution ending up outside of the field, in this case. The directive effect turns out to be not too strict and delayed compared to the classic FTLE. The crucial difference is that the compared work of Hlawatsch et al. obtains the FTLE field from the gradient $\nabla\mathbf{\upvarphi}(\mathbf{x})$ of the flow map, which is in turn obtained from sampling of trajectories following tensor field lines. We will now capture the same comparison for the textbook example \enquote{double gyre}.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[height=\textwidth]{img/gyre_ftle.png}
    \caption*{LTG($0^\circ$)}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_ftle_org.PNG}
     \caption*{FTLE: Hlawatsch's approach}
    \label{b)}
  \end{minipage}
\caption{\enquote{gyre} test field}
\label{gyre_ftle_comp}
\end{figure}

 We will now capture the star dataset revealing tensor field lines diverging in any direction from the center.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.29\textwidth}
  \includegraphics[height=\textwidth]{img/star.png}
    \caption*{\enquote{Star} test field}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.29\textwidth}
    \includegraphics[height=\textwidth]{img/star_ftle.PNG}
     \caption*{LTG slice($90^\circ$)}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.35\textwidth}
  \centering
    \includegraphics[height=\textwidth]{img/star_volume_alt.PNG}
     \caption*{IsoVolume on LTG volume}
    \label{b)}
  \end{minipage}
  \caption{LTG on \enquote{star} test field}
\label{star-ftle}
\end{figure}
When we look at the example depicted in Fig. \ref{star-ftle}, we can see the FTLE approach LTG responding most in the center of the field, which is the knot point of tensor field lines diverging in any direction. Remark, that there is a light V-shaped ridge on the left on blue background which is outshined by the center spot because of its orders of larger magnitude. The volume illustrated on the right even shows a spiked structure for the volume representing all directions at once. The approach detects more ridges in diagonal directions stochastically as we can infer from the Gaussian-shaped plates in the thresholded volume, evenly spaced in $90^\circ$ shifts with $45^\circ$ (quarter volume) offset. These can be interpreted nicely, since the diagonal direction is never offset into its transmission direction leading to a trivial (none) difference for axis-parallel directions for the offset direction orthogonal to the light source direction. Hence, the differences will be factor $2$ at worst for the diagonals, since we measure differences in both offset directions, in case ridges in both axis-parallel and -diagonal directions are present. Accordingly, we observe Gaussian-plates in the volume instead of sharp, contrasted ridges for the \enquote{star} example. But wouldn't we expect a ridges only in parallel directions? Well, in this case, we could imagine not only tensor field lines diverge or converge, but also they could cross each other unaffectedly (just as with light paths), which could be the case when shifting away from the diagonal directions. For better imagination, offset the light source vertically or horizontally from the diagonals and observe the light paths hitting the axes (, which in fact do cross each other).

@@

Now, we will evaluate a classic textbook example called the \enquote{Double Gyre} \cite{shawn}, which is a quadruple gyre in our case for formatting reasons, yet yielding the same conceptual results. We compare the glyph representation to the LTG magnitude for a single slice ($0^\circ$) again:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=\textwidth]{img/gyre.png}
    \caption{\enquote{Gyre} test field}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_ftle.PNG}
     \caption{LTG on \enquote{Gyre}($0^\circ$)}
    \label{b)}
  \end{minipage}
\label{drain_contour}
\end{figure}
We can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t. the $x$-axis ($0^\circ$). Note that even regions in antiparallel directions ridgres are detected to (literally) some degree: Remark the ridges being contiguos lines at the level of the gyre centers where converging turns into diverging. Also remember, that the resulting ridges are not to be concerned sharp ones, since we measure the separation of whole light distributions, which should be considered a continous response or measure. We again depict volume representations for this example:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
  \includegraphics[height=\textwidth]{img/gyre_volume2.png}
    \caption*{IsoVolume)}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_gauss_contour.PNG}
     \caption*{Contour}
    \label{b)}
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_volume4.PNG}
     \caption*{GaussContour}
    \label{b)}
  \end{minipage}
  \caption{\enquote{Gyre} test field}
\label{gyre_volume}
\end{figure}
When we consider the volumes for the \enquote{gyre} test field, we can remark the holes, which represent the gyres in the centers of the volume for a low and a high threshold, whereby we extract (filter out) the previously observed ridges in $3D$ for all directions $\omega$ simultaneously.

Now, that it passed the textbook examples, we push it forward to be applied in real (measured) data to show its performance in discerning and segmenting ROIs from a human brain. The approach will detect LCS in the brain's tissue from MRI-measured diffusion tensors of a male subject at the age of 27. We compare ParaView's glyph representation to the FTLE method LTG for the test field \enquote{brain}:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=\textwidth]{img/brain_cyl_glyphs.PNG}
    \caption*{glyph representation}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/brain_ftle.PNG}
     \caption*{LTG}
    \label{b)}
  \end{minipage}
  \caption{\enquote{brain} test field}
\label{brain-ftle}
\end{figure}
Regarding Fig. \ref{gyre_volume}, we find that the approach detects the boundaries of concavities (lat.: Gyri) in the brain, forming the lobus cerebri (commonly known a lobes). The invented method could thus ,e.g. , aid to compute $3D$ models of a subject in medical sciences. It emphasizes regions with much flow-rate (throughput), because in this areas the anisotropy and/or tensor magnitude is increased, preventing the intensities from being absorbed, hence yielding to a high FTLE-response. The regions highlighted in red, are caused by high linear anisotropy measures, as we can infer from the glyph representation. Note that we do not show the volume representation in this case, since all directions (slices) look approximately the same because the tensor field is only introducing changes in tensor magnitude not in anisotropy as inferred from the glyph representation. The regions with a tensor magnitude higher than zero restrict the light propagation to certain areas with the constraint, that not all of the energy is absorbed by sorrounding cells in the local neighborhood.

>>>>

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/tensorfieldlines.png}
    \label{a)}
    \caption*{a) TFLs and glyphs }
  \end{minipage}
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/rings_ftle.png}
    \label{b)}
    \caption*{b) LTG($0^\circ$) for a)}
  \end{minipage}
  \caption{``rings''-test field}
\label{rings-ftle}
\end{figure}
What we can observe in Fig. \ref{rings-ftle}, is called an inside-outside test producing the ridge on the circumference of the circle with maximized radius fitting into the minimum bounding box. If one would place a light source on the circumference of this circle offsetting it spatially and in the angular domain, one would obtain one incomplete final light distribution (outside) and one complete final light distribution (inside), as this is the maximum radius at which the intensities can follow a circle. This hence leads to large differential light distributions and thus to a higher LTG magnitude. ROIs pointing outside the field respond the same for varying light source directions and positions yielding the blue areas in the corners. Regions introducing linear anisotropy respond uniformly inside of the circle in contrary to the center point, which is again not responding because of circular overlaps of the offset light distributions computed for the light transport gradient. The contour filter reveals these two edge locations in the field, concerning tensor field line divergence.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/rings_contour.png}
    \caption*{Contour}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/rings_isovolume.png}
    \caption*{IsoVolume}
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/rings_volume.png}
    \caption*{Volume}
  \end{minipage}
    \caption{LTG volume representations: ``rings'' test field}
\label{rings_contour}
\end{figure}
The volumes, we examine here in Fig. \ref{rings_contour}, show that the field is symmetric, as we do not notice a change along the $\omega\mapsto z$-axis except for the natural shift of the field against the light source direction, which we discussed already in this subsection. Along the $z$-axis, the light source is rotated in all directions which causes the shift to form a full cycle from bottom to top ($[0^\circ,360^\circ]$).
At last, we will analyze the bow test field, which is now considered as a more complicated example. We again plot the FTLE field in color coding representation in Fig. \ref{bow-ftle}.

@@

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[height=\textwidth]{img/ftle_drain_alt.png}
    \caption*{LTG($0^\circ$)}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/drain_ftle_org.PNG}
     \caption*{FTLE: Hlawatsch's approach}
    \label{b)}
  \end{minipage}
\caption{\enquote{drain} test field}
\label{drain_ftle_comp}
\end{figure}
We can observe the comparison between our invented FTLE-like approach and the one of Hlawatsch et al. in the depiction in Fig. \ref{drain_ftle_comp}. Note, that a sharper but weaker ridge is detected on the right side of the image, which shines similar to a cone shape from the top. We interpret the missing spot in the top center as an artifact stemming from the way we handle our light propagation, as we allow whole distributions of trajectories to be propagated (with a scalar attribute interpretable as propability). This causes a sampling drift, we already discussed in sect. \ref{sect:total}, leading to an energy distribution ending up outside of the field, in this case. The directive effect turns out to be not too strict and delayed compared to the classic FTLE. The crucial difference is that the compared work of Hlawatsch et al. obtains the FTLE field from the gradient $\nabla\mathbf{\upvarphi}(\mathbf{x})$ of the flow map, which is in turn obtained from sampling of trajectories following tensor field lines. We will now capture the same comparison for the textbook example \enquote{double gyre}.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[height=\textwidth]{img/gyre_ftle.png}
    \caption*{LTG($0^\circ$)}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_ftle_org.PNG}
     \caption*{FTLE: Hlawatsch's approach}
    \label{b)}
  \end{minipage}
\caption{\enquote{gyre} test field}
\label{gyre_ftle_comp}
\end{figure}
Again, it is noticeable, that the results correspond to each other to a certain degree. This implies, that we obtained reasonable results generated by our method, but the classical FTLE is more strict concerning its directive abilities and congruently does not do things by halves (like we do with our partitioned light distributions). We will now capture the star dataset revealing tensor field lines diverging in any direction from the center.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.29\textwidth}
  \includegraphics[height=\textwidth]{img/star.png}
    \caption*{\enquote{Star} test field}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.29\textwidth}
    \includegraphics[height=\textwidth]{img/star_ftle.PNG}
     \caption*{LTG slice($90^\circ$)}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.35\textwidth}
  \centering
    \includegraphics[height=\textwidth]{img/star_volume_alt.PNG}
     \caption*{IsoVolume on LTG volume}
    \label{b)}
  \end{minipage}
  \caption{LTG on \enquote{star} test field}
\label{star-ftle}
\end{figure}
When we look at the example depicted in Fig. \ref{star-ftle}, we can see the FTLE approach LTG responding most in the center of the field, which is the knot point of tensor field lines diverging in any direction. Remark, that in the middle image there is a light V-shaped ridge on the left on blue background which is outshined by the center spot because of its orders of larger magnitude. The volume illustrated on the right even shows a spiked structure for the volume representing all directions at once. The approach detects more ridges in diagonal directions stochastically as we can infer from the Gaussian-shaped plates in the thresholded volume, evenly spaced in $90^\circ$ shifts with $45^\circ$ (quarter volume) offset. These can be interpreted nicely, since the diagonal direction is never offset into its transmission direction leading to a trivial (none) difference for axis-parallel directions for the offset direction orthogonal to the light source direction. Hence, the differences will be factor $2$ at worst for the diagonals, since we measure differences in both offset directions, in case ridges in both axis-parallel and -diagonal directions are present. Accordingly, we observe Gaussian-plates in the volume instead of sharp, contrasted ridges for the \enquote{star} example. But wouldn't we expect a ridges only in parallel directions? Well, in this case, we could imagine not only tensor field lines diverge or converge, but also they could cross each other unaffectedly (just as with light paths), which could be the case when shifting away from the diagonal directions. For better imagination, offset the light source vertically or horizontally from the diagonals and observe the light paths hitting the axes (, which in fact do cross each other).

@@

Now, we will evaluate a classic textbook example called the \enquote{double gyre} \cite{shawn}, which is a quadruple gyre in our case for formatting reasons, yet yielding the same conceptual results. We compare the glyph representation to the LTG magnitude for a single slice ($0^\circ$) again:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=\textwidth]{img/gyre.png}
    \caption{\enquote{Gyre} test field}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_ftle.PNG}
     \caption{LTG on \enquote{Gyre}($0^\circ$)}
    \label{b)}
  \end{minipage}
\label{drain_contour}
\end{figure}
We can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t. the $x$-axis ($0^\circ$). Note that even regions in antiparallel directions ridgres are detected to (literally) some degree: Remark the ridges being contiguos lines at the level of the gyre centers where converging turns into diverging. Also remember, that the resulting ridges are not to be concerned sharp ones, since we measure the separation of whole light distributions, which should be considered a continous response or measure. We again depict volume representations for this example:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
  \includegraphics[height=\textwidth]{img/gyre_volume2.png}
    \caption*{IsoVolume)}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_gauss_contour.PNG}
     \caption*{Contour}
    \label{b)}
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_volume4.PNG}
     \caption*{GaussContour}
    \label{b)}
  \end{minipage}
  \caption{\enquote{Gyre} test field}
\label{gyre_volume}
\end{figure}
When we consider the volumes for the \enquote{gyre} test field, we can remark the holes, which represent the gyres in the centers of the volume for a low and a high threshold, whereby we extract (filter out) the previously observed ridges in $3D$ for all directions $\omega$ simultaneously.

Now, that the approach has passed the textbook examples, we push it forward to be applied in real (measured) data to show its performance in discerning and segmenting ROIs from a human brain. The approach will detect LCS in the brain's tissue from MRI-measured diffusion tensors of a male subject at the age of 27. We compare ParaView's glyph representation to the FTLE method LTG for the test field \enquote{brain}:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=\textwidth]{img/brain_cyl_glyphs.PNG}
    \caption*{glyph representation}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/brain_ftle.PNG}
     \caption*{LTG}
    \label{b)}
  \end{minipage}
  \caption{\enquote{brain} test field}
\label{brain-ftle}
\end{figure}
Regarding Fig. \ref{gyre_volume}, we find that the approach detects the boundaries of concavities (lat.: Gyri) in the brain, forming the lobus cerebri (commonly known a lobes). The invented method could thus ,e.g. , aid to compute $3D$ models of a subject in medical sciences. It emphasizes regions with much flow-rate (throughput), because in this areas the anisotropy and/or tensor magnitude is increased, preventing the intensities from being absorbed, hence yielding to a high FTLE-response. The regions highlighted in red, are caused by high linear anisotropy measures, as we can infer from the glyph representation. Note that we do not show the volume representation in this case, since all directions (slices) look approximately the same because the tensor field is only introducing changes in tensor magnitude not in anisotropy as inferred from the glyph representation. The regions with a tensor magnitude higher than zero restrict the light propagation to certain areas with the constraint, that not all of the energy is absorbed by sorrounding cells in the local neighborhood.
